---
title: "habitual priors"
author: "mht"
date: "November 18, 2015"
output: html_document
---


To reconstruct a prevalence-like distribution from the frequency dependent measure, 
we will assume there is a maximum frequency (akin to doing the action "all the time").
This will probably be set by the item: *smokes*, because it is the thing that could happen the most. 


# Prior elicitation 1

See priors1-process.Rmd for details on what was done to the data up to thsi point.

## Data analysis

```{r}
d<-read.csv("~/Documents/research/habituals/data/priors/priors-1.csv")
```

First let's look at the slider bar question (what % of the US population has done this action before?)

```{r}
## histogram of responses for existence questions
qplot(data=d, x = existence, geom='histogram')
```

Some responses are that the action is super common, or super rare, or somewhere in between. Nice spread.

Second, let's look at the times per X question.

```{r}
qplot(data=d, x = nTimes, geom='histogram')+
  facet_wrap(~timeWindow, scales='free')


range(filter(d,timeWindow=='day')$nTimes)
range(filter(d,timeWindow=='life')$nTimes)
```

Looks like the most common frequent thing happens (response) 20 times / day and the least common thing happens (response) is once (or 0 times) in a life time.

That's kind of strange. The question asked to assume that the person did it at least once. I guess there is an interpretation that the question is asking how many *more times* will they do it? How many people said 0 times in a life?

```{r}
table(filter(d,timeWindow=='life')$nTimes)
```

Just two responses were like this. All of the others said once in a life time. 

Let's scale the data so that we're all the responses are on the same scale.
We'll see that a life time = 80 years, and scale everything to the year scale.


```{r}
scaleTime <- list(
  day = 365,
  week = 52,
  month = 12,
  year = 1,
  life = 1/80
)

d.tidy<-d %>% 
  mutate(timeWindow = as.character(timeWindow)) %>%
  rowwise() %>%
  mutate(frequency = nTimes*scaleTime[[timeWindow]]) %>%
  ungroup() 

ggplot(d.tidy, aes(x=frequency))+
  geom_histogram()+
  facet_wrap(~item, scales='free')

d.stats<-left_join(
  d.tidy %>% 
  group_by(item) %>%
  multi_boot_standard(column="frequency") %>%
  rename(freq = mean,
         freqHi = ci_upper,
         freqLo = ci_lower),
  d.tidy %>% 
  group_by(item) %>%
  multi_boot_standard(column="existence") %>%
  rename(exist = mean,
         existHi = ci_upper,
         existLo = ci_lower))

ggplot(d.stats, aes(x=freq, y = exist, color=item))+
  geom_point()+
  geom_errorbarh(aes(xmin=freqLo, xmax=freqHi))+
  geom_errorbar(aes(ymin=existLo, ymax=existHi))+
  xlim(-0.05,1000)+
  ylim(-0.05,1.05)#+
  #coord_fixed(ratio=1)


```