---
title: "habituals priors process"
author: "mht"
date: "November 18, 2015"
output: html_document
---


## Data processing on priors-1

I accidentally collected 5 subjects as pilot (which according to MTurk rules means you can't collect more than 9 in that batch).
So I ran a second version with 25 subjects.

Did any subject take it twice?
```{r duplicate.subjects}
s1<- read.csv("~/Documents/research/habituals/mturk/priors-1/priors-1_invoice.csv")[1:5,]
s2<- read.csv("~/Documents/research/habituals/mturk/priors-1b/priors-1b_invoice.csv")[1:25,]
s3<-bind_rows(s1,s2)
length(levels(factor(s3$workerid)))
```
Phew. No one took it twice. Have a good sample of n=30.

Note: Drop down menu check didn't work. However, all but 3 trials (out of 20*30 = 600) had participants respond with the menu.
One of the subjects reported that the dropdown menu didn't work for his last trial, and recorded his response in the comments section.

Load first 5 participants data.

```{r load.d1}
d1<- read.csv("~/Documents/research/habituals/mturk/priors-1/priors-1-trials.csv")
comments1<- read.csv("~/Documents/research/habituals/mturk/priors-1/priors-1-subject_information.csv")
print(comments1[[3,"problems"]])
# modify participant's data to reflect his comments
d1[60, 5]<- "week"

# comprehension question information
catch1<- read.csv("~/Documents/research/habituals/mturk/priors-1/priors-1-catch_trials.csv")
# everyone passed
```

Load second group (n=25) data and merge 2 data sets.
```{r load.d2}
d2<- read.csv("~/Documents/research/habituals/mturk/priors-1b/priors-1b-trials.csv")

# comprehension question information
catch2<- read.csv("~/Documents/research/habituals/mturk/priors-1b/priors-1b-catch_trials.csv")
failCatch<-filter(catch2, pass==0)$workerid
d2 <- d2 %>% 
  filter(!(workerid == failCatch))

# renumber workerids
d2$workerid <- d2$workerid+max(d1$workerid) + 1
d<- bind_rows(d1,d2)

# remove trials with no dropdown menu information

d<-d %>%
  filter(!(is.na(timeWindow)))

## write to csv so you don't have to do this everytime
#write.csv(d, file="~/Documents/research/habituals/data/priors/priors-1.csv", row.names=F)
```
